Nengo: A Python tool for building large-scale functional brain models

### 2013-10-02 nengo.tex
Fixed up figures, esp. NEF
Figures done

\def\journal{Neuroinformatics}
\def\DOI{}
\def\articleType{Methods Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Bekolay {et~al.}}
\def\Authors{Trevor Bekolay\,$^{1,*}$, James Bergstra\,$^{1}$,
  Xuan Choo\,$^{1}$, Travis DeWolf\,$^{1}$, Eric Hunsberger\,$^{1}$,
  Daniel Rasmussen\,$^{1}$, Terrence C. Stewart\,$^{1}$
  and Chris Eliasmith\,$^{1}$}
\def\Address{$^{1}$Computational Neuroscience Research Group,
  Center for Theoretical Neuroscience,
  University of Waterloo, Waterloo, ON, Canada}
\def\corrAuthor{Trevor Bekolay}
\def\corrAddress{David R. Cheriton School of Computer Science,
  University of Waterloo, 200 University Avenue West,
  Waterloo, ON, N2L 3G1, Canada}
\def\corrEmail{tbekolay@uwaterloo.ca}

@@

\begin{abstract}
%Maximum: 2000 characters

  \tiny \keyFont{\section{Keywords:} 1 2 3 4 5 (6 7 8)}
\end{abstract}

@@

\begin{table}[!t]
\processtable{Nengo tools\label{tab:nengo-vers}}
{\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
\textbf{Name} & \textbf{Implementation language}
  & \textbf{Code repository} & \textbf{Purpose} \\\midrule
JavaNengo simulator & Java & \texttt{nengo\_java}
  & Simulates many types of spiking neuron models.
  Includes methods to construct and simulate models using the NEF. \\
JavaNengo GUI & Java & \texttt{nengo\_java\_gui}
  & Graphical interface to JavaNengo simulator.
  Allows modellers to drag-and-drop components built with the NEF. \\
JavaNengo scripting interface & Jython & \texttt{nengo\_jython}
  & Scripting interface to JavaNengo simulator.
  Provides shortcuts to classes and functions in the simulator.
  Also accessible through and interacts with the JavaNengo GUI. \\
PyNengo & Python & \texttt{nengo}
  & Constructs and simulates models using the NEF. \\
PyNengo OpenCL simulator & Python+PyOpenCL
  & \texttt{nengo\_ocl}
  & Simulates models constructed with PyNengo. \\\botrule
\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}}
\end{table}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.19\textwidth]{nef_summary_enc}
  \includegraphics[width=0.19\textwidth]{nef_summary_dec}
  \includegraphics[width=0.399\textwidth]{nef_summary_trans}
  \includegraphics[width=0.199\textwidth]{nef_summary_dyn}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
   Summary of the Neural Engineering Framework.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.6\textwidth]{sim}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
   The first two timesteps of a simple simulation.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=.9\textwidth]{comm_channel}
  \begin{minipage}{0.58\textwidth}
    \begin{lstlisting}
      import nengo
      from nengo.helpers import white_noise
      model = nengo.Model("Communication Channel")
      model.make_node("Input", white_noise(1, 5))
      model.make_ensemble("A", nengo.LIF(30), 1)
      model.make_ensemble("B", nengo.LIF(30), 1)
      model.connect("Input", "A")
      model.connect("A", "B")
      model.probe("Input")
      model.probe("A", filter=0.01)
      model.probe("A.spikes")
      model.probe("B", filter=0.01)
      model.probe("B.spikes")
      sim = model.simulator()
      sim.run(1)
      A_data = sim.data("A")
      A_spikes = sim.data("A.spikes")
      ...
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.37\textwidth}
    \includegraphics[width=\textwidth]{comm_channel_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:comm-channel}
   Figure \arabic{figure}.}{
   Communcation channel model. (Top) Diagram of the model.
   (Bottom left) Python code to create and run the model.
   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \begin{minipage}{.495\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      # We use Nengo to generate
      #     bias: Bias currents
      #     decoders: Decoding weights
      #     encoders: Encoding weights
      # for populations A and B.
      import pynn.nest as pyNN

      lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
                     'tau_syn_I':100}
      dt = 0.5
      pyNN.setup(timestep=dt, min_delay=dt)
      A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
      B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
      input_signal = 0.5
      inputA = [pyNN.DCSource(amplitude=val) for val
                 in A_bias + input_signal * A_encoders]
      for i, pulse in enumerate(inputA):
          pulse.inject_into(A[i:i+1])
      inputB = [pyNN.DCSource(amplitude=val) for val
                 in B_bias]
      for i, pulse in enumerate(inputB):
          pulse.inject_into(B[i:i+1])
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{.495\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      weights = []
      for i in xrange(30):
          for j in xrange(30):
              weights.append((i, j,
                  np.dot(A_decoder[i], B_encoder[j]), 1.0))
      connection = pyNN.Projection(
          A, B, pyNN.FromListConnector(weights))
      A.record('spikes')
      B.record('spikes')
      pyNN.run(1000)

      # Decoded value of B
      B_output = numpy.zeros(1000)
      for i in xrange(1000):
          spikes = B[i:i+1].getSpikes()[:,1]
          spikes = (spikes / dt).astype('int')
          B_output[spikes] += B_decoder[i]
      # Filter with tau = 200 ms
      decay = np.exp(-dt / 200)
      B_output[0, :] *= (1 - decay)
      for i in xrange(1, 1000):
          B_output[i,:] = decay * B_output[i-1,:]
                           + (1-decay) * B_output[i,:]
    \end{lstlisting}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
   An implementation of the communication channel in PyNN.}
\end{figure}

@@

      model.connect('State', 'State',
                      function=feedback, filter=tau)
      model.probe('State', filter=tau)
      model.probe('State.spikes')
      sim = model.simulator()
      sim.run(6)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.55\textwidth}
    \includegraphics[width=0.95\textwidth]{lorenz}
    \includegraphics[width=\textwidth]{lorenz_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
   Lorenz attractor model. (Left) Python code to create and run the model.
   (Top right) Diagram of the model.
   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.8\textwidth]{cconv}
  \begin{minipage}{0.53\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      import nengo
      from nengo.networks import CircularConvolution
      model = nengo.Model("Circular convolution")
      model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
      model.make_ensemble("A", nengo.LIF(512), 4)
      model.connect("Input A", "A")
      model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
      model.make_ensemble("B", nengo.LIF(512), 4)
      model.connect("Input B", "B")
      cconv = model.add(CircularConvolution("Circular Convolution",
          neurons=nengo.LIF(1032), dimensions=4))
      model.connect("A", cconv.A)
      model.connect("B", cconv.B)
      model.make_ensemble("Result", nengo.LIF(512), 4)
      model.connect(cconv, "Result")
      model.probe("A", filter=0.02)
      model.probe("B", filter=0.02)
      model.probe("Result", filter=0.02)
      sim = model.simulator()
      sim.run(0.5)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.43\textwidth}
    \includegraphics[width=\textwidth]{cconv_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:cconv} Figure \arabic{figure}.}{
   Circular convolution model. (Top) Diagram of the model.
   Note that much of the complexity is contained within the
   circular convolution network; the transform matrices shown
   are automatically generated by the network depending on
   input dimensionality.
   (Bottom left) Python code to create and run the model.

   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.5\textwidth]{ocl_vs_ref}
\end{center}
 \textbf{
     \refstepcounter{figure}\label{fig:ocl_vs_ref}
     Figure \arabic{figure}.}{
     Five-hundred-dimensional circular convolution networks were some of the most
     computationally demanding elements of the Spaun cognitive architecture.
     Simulations of circular convolution networks of various sizes illustrate the speed of our re-engineered simulators.
     A straightforward reference implementation in Python is sufficient for small networks,
     while for larger ones the new OpenCL-powered simulator runs up to 200$\times$ faster,
     In particular the Radeon 7970 GPU can simulate 500-dimensional convolution networks faster than real time (0.95 seconds for a 1-second simulation).
     Also, although both are multithreaded, our new implementation delivers better CPU performance than the previous JavaNengo implementation.
     }
\end{figure}

>>>>

\def\journal{Neuroinformatics}
\def\DOI{}
\def\articleType{Methods Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Bekolay {et~al.}}
\def\Authors{Trevor Bekolay\,$^{1,*}$, James Bergstra\,$^{1}$,
  Eric Hunsberger\,$^{1}$, Travis DeWolf\,$^{1}$, Terrence C. Stewart\,$^{1}$,
  Daniel Rasmussen\,$^{1}$, Xuan Choo\,$^{1}$, Aaron Voelker\,$^{1}$,
  and Chris Eliasmith\,$^{1}$}
\def\Address{$^{1}$Computational Neuroscience Research Group,
  Center for Theoretical Neuroscience,
  University of Waterloo, Waterloo, ON, Canada}
\def\corrAuthor{Trevor Bekolay}
\def\corrAddress{David R. Cheriton School of Computer Science,
  University of Waterloo, 200 University Avenue West,
  Waterloo, ON, N2L 3G1, Canada}
\def\corrEmail{tbekolay@uwaterloo.ca}

@@

\begin{abstract}
  Neuroscience currently lacks
  a comprehensive theory of
  how the brain implements cognitive processes
  with biological parts.
  The Neural Engineering Framework (NEF)
  proposes one such theory,
  but has not has not gathered
  significant empirical support
  due, in part, to the technical
  challenge of building and simulating
  large-scale models with the NEF.
  Nengo is a tool that builds and simulates
  functional brain models with the NEF.
  Nengo is the primary resource
  for teaching how the NEF
  is used to model neural systems,
  and for researchers using the NEF
  to build and simulate
  functional brain models.
  A previous version of Nengo was used
  to create Spaun,
  the world's largest functional brain model.
  The implementation of Spaun
  identified limitations
  in Nengo's ability to
  construct models with simple syntax,
  simulate them quickly,
  and collect data for subsequent analysis.
  We have implemented a new version of Nengo
  in Python that overcomes these limitations.
  It uses simple and extendable syntax,
  simulates a benchmark model on the scale of Spaun
  200 times faster than the previous implementation,
  and has a flexible mechanism
  for collecting simulation results.

  \tiny \keyFont{\section{Keywords:} Python, Neural Engineering
    Framework, theoretical neuroscience, computational neuroscience,
    control theory, software, simulation}
\end{abstract}

@@

\begin{table}[!t]
\processtable{Tools bearing the name Nengo.\label{tab:nengo-vers}}
{\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
\textbf{Name} & \textbf{Implementation language}
  & \textbf{Code repository} & \textbf{Purpose} \\\midrule
JavaNengo simulator & Java & \texttt{nengo\_java}
  & Simulates many types of spiking neuron models.
  Includes methods to construct and simulate models using the NEF. \\
JavaNengo GUI & Java & \texttt{nengo\_java\_gui}
  & Graphical interface to JavaNengo simulator.
  Allows modellers to drag-and-drop components built with the NEF. \\
JavaNengo scripting interface & Jython & \texttt{nengo\_jython}
  & Scripting interface to JavaNengo simulator.
  Provides shortcuts to classes and functions in the simulator.
  Also accessible through and interacts with the JavaNengo GUI. \\
PyNengo & Python & \texttt{nengo}
  & Constructs and simulates models using the NEF. \\
PyNengo OpenCL simulator & Python+PyOpenCL
  & \texttt{nengo\_ocl}
  & Simulates models constructed with PyNengo. \\\botrule
\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}.}
\end{table}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.245\textwidth]{nef_summary_enc}
  \includegraphics[width=0.245\textwidth]{nef_summary_dec}
  \includegraphics[width=0.245\textwidth]{nef_summary_trans}
  \includegraphics[width=0.245\textwidth]{nef_summary_dyn}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
   Summary of the three principles of the Neural Engineering Framework
   (NEF). (A) By the representation principle, signals are encoded
   in neural populations based on the \textit{tuning curve}
   of each neuron (top). The tuning curve describes
   how active a neuron is given some input signal.
   If we drive the eight neurons in the top panel
   with the input signal in the middle panel,
   we see the firing pattern in the bottom panel.
   (B) By the representation principle,
   the spiking activity of a neural population
   can be decoded to recover the original input signal,
   or some transformation of that input signal.
   First, the firing pattern shown in the top panel
   is filtered with a decaying exponential filter (middle panel).
   The filtered activity is then summed together
   with a set of weights that approximates
   the input signal (bottom panel, green)
   and the cosine of the input signal (bottom panel, purple).
   (C) By the transformation principle,
   populations of neurons can send signals
   to another population by decoding
   the desired function from the first population
   and then encoding the decoded estimate
   into the second population.
   These two steps can be expanded into a single step
   by calculating a set of weights
   that describe the strength of the connection
   between each neuron in the first population
   and each neuron in the second population.
   In (C), a sine wave is encoded by population A (top panel);
   the negative of that signal is projected
   to population B (middle panel)
   and the square of that signal is projected
   to population C (bottom panel).
   (D) By the dynamcs principle,
   signals being represented by population of neurons
   can be thought of
   as state variables in a dynamical system.
   In this example, the dynamical system
   has negative feedback across its two dimensions,
   resulting in a harmonic oscillator.
   The oscillator is plotted across time (top)
   and in state space (bottom).}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=\textwidth]{sim}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
   Detailed breakdown of the PyNengo reference simulator
   running a simple model for two timesteps.
   (A) Code describing the model being simulated.
   It consists of ensemble A project the sine of its
   encoded signal to ensemble B,
   which is recurrently connected.
   (B) Diagram depicting the model being simulated.
   (C) A detailed diagram of how the reference simulator
   simulates this model. Signals, which are the values
   tracked in the simulation, are colored blue.
   Operators, which are the computations done on signals,
   are colored red.
   Signals can be grouped as low-level neural signals
   that are used to compute the nonlinear functions
   underlying neuron models,
   and high-level NEF signals that are used to
   drive neurons and track the signals
   that the neurons are representing.
   The operators that implement the decoding
   and encoding steps map between
   the low-level neural signals
   and the high-level NEF signals.
   (D) The signals tracked at the low level
   can be interpreted as a model
   commonly seen in computational neuroscience literature;
   a population of leaky integrate-and-fire neurons
   is driven by some time-varying input current, $J(t)$.
   These neurons project to a population
   of recurrently connected neurons.
   The connection weight between the two populations,
   and from the second population to itself,
   can be computed by the NEF's transformation
   principle, bypassing the need for
   the high-level NEF signals
   used by the reference simulator
   for speed and data collection purposes.
   (E) The signals tracked at the high level
   can be interpreted as a dynamical system
   that can be analyzed with control theory.
   A simply represents its input,
   and passes its state to a sin function
   which becomes the input to B.
   B is a simple linear system
   that can be described with the typical
   $\dot{x}(t) = A x(t) + B u(t)$ equation.
   These dynamical systems can be simulated
   directly, without the use of spiking neurons,
   in order to quickly analyze system behavior.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=.85\textwidth]{comm_channel}
  \begin{minipage}{0.48\textwidth}
    \begin{lstlisting}
import nengo
from nengo.helpers import white_noise
model = nengo.Model("Communication Channel")
model.make_node("Input", white_noise(1, 5))
model.make_ensemble("A", nengo.LIF(30), 1)
model.make_ensemble("B", nengo.LIF(30), 1)
model.connect("Input", "A")
model.connect("A", "B")
model.probe("Input")
model.probe("A", filter=0.01)
model.probe("A.spikes")
model.probe("B", filter=0.01)
model.probe("B.spikes")
sim = model.simulator()
sim.run(1)
A_data = sim.data("A")
A_spikes = sim.data("A.spikes")
...
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.39\textwidth}
    \includegraphics[width=\textwidth]{comm_channel_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:comm-channel}
   Figure \arabic{figure}.}{
   A communication channel implemented with Nengo.
   (A) Diagram depicting the model. Ensemble A
   projects its encoded signal to ensemble B unchanged.
   (B) Nengo code to build and simulate the model
   for 1 second.
   (C) The results of the simulation.
   The input signal (top panel) is white noise limited to 0 to 5 Hz.
   The signal is well represented by both ensemble A (middle panel)
   and ensemble B (bottom panel) despite the neural firing patterns
   (underlaid in middle and bottom panels) being different.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \begin{minipage}{.435\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
# We use Nengo to generate
#     bias: Bias currents
#     decoders: Decoding weights
#     encoders: Encoding weights
# for populations A and B.
import pynn.nest as pyNN

lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
               'tau_syn_I':100}
dt = 0.5
pyNN.setup(timestep=dt, min_delay=dt)
A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
input_signal = 0.5
inputA = [pyNN.DCSource(amplitude=val) for val
           in A_bias + input_signal * A_encoders]
for i, pulse in enumerate(inputA):
    pulse.inject_into(A[i:i+1])
inputB = [pyNN.DCSource(amplitude=val) for val
           in B_bias]
for i, pulse in enumerate(inputB):
    pulse.inject_into(B[i:i+1])
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{.46\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
weights = []
for i in xrange(30):
    for j in xrange(30):
        weights.append((i, j,
            np.dot(A_decoder[i], B_encoder[j]), 1.0))
connection = pyNN.Projection(
    A, B, pyNN.FromListConnector(weights))
A.record('spikes')
B.record('spikes')
pyNN.run(1000)

# Decoded value of B
B_output = numpy.zeros(1000)
for i in xrange(1000):
    spikes = B[i:i+1].getSpikes()[:,1]
    spikes = (spikes / dt).astype('int')
    B_output[spikes] += B_decoder[i]
# Filter with tau = 200 ms
decay = np.exp(-dt / 200)
B_output[0, :] *= (1 - decay)
for i in xrange(1, 1000):
    B_output[i,:] = decay * B_output[i-1,:]
                     + (1-decay) * B_output[i,:]
    \end{lstlisting}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
   Implementation of the communication channel in PyNN.
   Note that several quantities are generated from Nengo;
   the functions that generate these quantities in Nengo
   can be decoupled, but are not included here for brevity.
   The implementation of the Lorenz attractor
   is not shown, but is very similar to the communication channel.}
\end{figure}

@@

model.connect('State', 'State',
                function=feedback, filter=tau)
model.probe('State', filter=tau)
model.probe('State.spikes')
sim = model.simulator()
sim.run(6)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.55\textwidth}
    \includegraphics[width=0.95\textwidth]{lorenz}
    \includegraphics[width=\textwidth]{lorenz_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
   A Lorenz attractor implemented with Nengo.
   (A) Nengo code to build and simulate the model
   for 6 seconds.
   (B) Diagram depicting the model. The state ensemble
   is recurrently connected with complex dynamics
   implemented the chaotic Lorenz attractor.
   Not that this populations does not receive
   any input that might drive its initial value;
   instead, the initial value is determined by
   the baseline firing of the 2000 leaky integrate-and-fire
   neurons that make up the state ensemble.
   (C) The trajectory that the state ensemble takes
   in its three-dimensional state space.
   For the parameters chosen, the trajectory takes
   the well-known butterfly shape.
   (D) The state vector plotted over time.
   (E) The spikes emitted by a random sample of 25
   neurons from the state ensemble.
   Some neurons fire uniformly across the 6 second simulation,
   but most change depending on the state being tracked
   due to the recurrent connection.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.8\textwidth]{cconv}
  \begin{minipage}{0.46\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
import nengo
from nengo.networks import CircularConvolution
model = nengo.Model("Circular convolution")
model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
model.make_ensemble("A", nengo.LIF(512), 4)
model.connect("Input A", "A")
model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
model.make_ensemble("B", nengo.LIF(512), 4)
model.connect("Input B", "B")
cconv = model.add(CircularConvolution("Circular Convolution",
    neurons=nengo.LIF(1032), dimensions=4))
model.connect("A", cconv.A)
model.connect("B", cconv.B)
model.make_ensemble("Result", nengo.LIF(512), 4)
model.connect(cconv, "Result")
model.probe("A", filter=0.02)
model.probe("B", filter=0.02)
model.probe("Result", filter=0.02)
sim = model.simulator()
sim.run(0.5)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.46\textwidth}
    \includegraphics[width=\textwidth]{cconv_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:cconv} Figure
   \arabic{figure}.}{
   Circular convolution implemented with Nengo.
   (A) Diagram depicting the model.
   The input vectors, A and B, represent four-dimensional signal
   which are mapped onto six ensembles within the
   Circular Convolution network through
   a complicated transformation matrix.
   Each ensemble within the network represents a
   two-dimensional signal. The product of those two signals
   is projected through another complicated transformation matrix
   to compute the final four-dimensional result.
   Note that the complicated parts of the model
   are contained within the network;
   the number of ensembles and the transform matrices shown
   are automatically generated by the network depending on
   the dimensionality of the input vectors.
   (B) Nengo code to build and simulate the model
   for 0.5 seconds.
   (C) The result of the simulation.
   Straight horizontal lines represent
   the target values that each ensemble
   should represent.
   Wavy lines represent the decoded values
   for each dimension represented by the
   A, B, and Result ensembles
   (top, middle, bottom panels respectively).
   The ensembles represent the correct values,
   after a startup transient of less than 0.1 seconds.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.32\textwidth]{bench_cchannel}
  \includegraphics[width=0.32\textwidth]{bench_lorenz}
  \includegraphics[width=0.32\textwidth]{bench_cconv}
\end{center}
 \textbf{
     \refstepcounter{figure}\label{fig:benchmarks}
     Figure \arabic{figure}.}{
     For (A) and (B), all of the simulators
     except the PyNengo OpenCL simulator
     was run on an Intel Core i7-965.
     JavaNengo used all 4 cores of this processor;
     all other simulator used only 1 core.
     The PyNengo OpenCL simulator
     was run on an NVidia GTX280 GPU.
     For (C), if not noted, the simulator used
     was the PyNengo OpenCL simulator.
     The CPU used for JavaNengo and
     the PyNengo reference simulator
     was an Intel Core i7-3770;
     all 4 cores were used by JavaNengo,
     while PyNengo's reference simulator
     only used one core.
     }
\end{figure}

While benchmark models are not indicative
of performance on all models,
increasing simulation speed
was a primary motivators in creating PyNengo.
To validate that performance has improved,
we ran the models described in section~\ref{sec:examples}
for various numbers of neurons and dimensions
for each ensemble.

The communication channel and Lorenz attractor
are small models that demonstrate
the principles of the NEF.
Their small size enables us to write
PyNN scripts that implement roughly
the same functionality
with Brian, NEURON, NEST, and TODO PCSim.
We run each parameter set five times
on the same machine,
and plot the mean.
In most cases, the coefficient of variation
is well below 0.1, except for two
outliers with coefficients of 0.18 and 0.22,
indicating that these means are robust.
The results, shown in Figure~\ref{fig:benchmarks},
suggest that all versions of Nengo are significantly
faster than the simulators accessible
through PyNN, especially
as the size of the model increases.
This is likely due to Nengo's
use of factorized weight matrices,
rather than storing and computing with
the entire weight matrix
on each timestep.
While NEST and NEURON were not
run with MPI,
the reference simulator of PyNengo
also only uses one CPU core.
The results also suggest that PyNengo's
OpenCL simulator is faster
than the reference simulator
and JavaNengo's simulator,
though the difference
is not overwhelming in these models.

As a real-world example,
we have also benchmarked
the circular convolution model.
Circular convolution is an important test case,
as a significant portion of Spaun's
2.5 million neurons are used to
implement circular convolution.
In this case, only versions of Nengo
were tested, and were tested rigorously.
Instead of running each simulation multiple times,
we instead run the simulator for 10 timesteps
in order to fill various levels of the cache,
and then run the simulator for 1000 more timesteps;
there is very little variance using this method.
As can be seen in Figure~\ref{fig:benchmarks},
for large models, the OpenCL simulator
performs much faster than JavaNengo;
in particular, a Radeon 7970 GPU performs 500-dimensional
circular convolution faster than real time,
and 200 times faster than JavaNengo.
Additionally, although both JavaNengo
and the OpenCL simulator on a CPU
both use all available CPU cores,
PyNengo's OpenCL simulator is significantly faster.
